---
title: "How does music change over time?"
description: |
  Visualize the differing trends in volume, tempo, energy, and popularity over time.
author:
  - name: Wanjia Guo
    url: https://wanjiag.github.io/
date: 03-05-2021
categories:
  - visualization
  - spotify
output:
  distill::distill_article:
    self_contained: false
---
```{r setup, include=FALSE}

library(tidyverse)
#remotes::install_github("wilkelab/cowplot")
#install.packages("colorspace", repos = "http://R-Forge.R-project.org")
#remotes::install_github("clauswilke/colorblindr")
library(colorblindr)
library(gghighlight)
library(ggtext)

theme_set(theme_minimal())

files <- fs::dir_ls(here::here("data"))

obs <- gsub(".+data/(.+)\\.csv", "\\1", files)
purrr::map2(obs, files, ~{
   assign(.x, read_csv(.y), envir = .GlobalEnv)
})
```

# Data Cleaning

```{r}

data1 = data %>% 
  select(id, popularity, year, danceability, duration_ms, tempo, speechiness) %>% 
  mutate(popularity = (popularity-min(popularity))/(max(popularity)-min(popularity)),
         danceability = (danceability-min(danceability))/(max(danceability)-min(danceability)),
         duration = 
           (duration_ms-min(duration_ms))/(max(duration_ms)-min(duration_ms)),
         tempo = (tempo-min(tempo))/(max(tempo)-min(tempo)),
         speechiness = (speechiness-min(speechiness))/(max(speechiness)-min(speechiness))) %>% 
  select(-duration_ms)
  

fig1_data = data1 %>% select(-year) %>% 
  mutate(danceability = cut(danceability, 100, labels=FALSE),
         duration = cut(duration, 100, labels=FALSE),
         tempo = cut(tempo, 100, labels=FALSE),
         speechiness = cut(speechiness, 100, labels=FALSE)) %>% 
  pivot_longer(cols = danceability:duration, 
               names_to = "property") %>% 
  group_by(property, value) %>% 
  summarise(popularity_median = median(popularity)) %>% 
  mutate(property = factor(property, levels=c("danceability",
                                              "tempo",
                                              "speechiness",
                                              "duration"),
                           labels = c("Danceability",
                                              "Tempo",
                                              "Speechiness",
                                              "Duration")))

fig2_data = data1 %>% 
  mutate(decades = year - year %% 10 ) %>% 
  select(-c(popularity, year)) %>% 
  pivot_longer(cols = danceability:duration, 
               names_to = "property") %>% 
  group_by(property, decades) %>% 
  summarise(mean = mean(value)) %>% 
  mutate(property = factor(property, levels=c("duration",
                                              "speechiness",
                                              "tempo",
                                              "danceability")))
```

```{r}

fig2_data2 = data_by_year %>% 
  select(-c(key, mode)) %>% 
  mutate(acousticness = (acousticness-min(acousticness))/(max(acousticness)-min(acousticness)),
         danceability = (danceability-min(danceability))/(max(danceability)-min(danceability)),
         duration = (duration_ms-min(duration_ms))/(max(duration_ms)-min(duration_ms)),
         energy = (energy-min(energy))/(max(energy)-min(energy)),
         instrumentalness =
           (instrumentalness-min(instrumentalness))/(max(instrumentalness)-min(instrumentalness)),
         liveness = (liveness-min(liveness))/(max(liveness)-min(liveness)),
         loudness = (loudness-min(loudness))/(max(loudness)-min(loudness)),
         speechiness = (speechiness-min(speechiness))/(max(speechiness)-min(speechiness)),
         tempo = (tempo-min(tempo))/(max(tempo)-min(tempo)),
         valence = (valence-min(valence))/(max(valence)-min(valence)),
         popularity = (popularity-min(popularity))/(max(popularity)-min(popularity))
         ) %>% 
  select(-duration_ms) %>% 
  pivot_longer(cols = acousticness:duration, 
               names_to = "property")

fig2_data3 = fig2_data2 %>% 
  filter(property %in% c('energy', 'loudness', 'tempo', 'popularity')) %>% 
  mutate(property = factor(property, levels=c('popularity','energy', 'loudness', 'tempo')))
```

# Final Visualization:

```{r layout = "l-body-outset"}
ggplot(fig2_data3 %>% 
         filter(property!="popularity"),
       aes(x=year, y=value)) + 
  geom_line(aes(color=property)) +
  geom_line(data=fig2_data3 %>% 
              filter(property=="popularity"), 
            aes(x=year, y=value), 
            color="red",
            size=1.3)+
  gghighlight(use_direct_label = FALSE)+
  scale_color_OkabeIto(darken=-0.5)+
  theme(plot.title = ggtext::element_markdown(),
        legend.position = "none")+
  labs(title = "Music is 
  <span style = 'color: #57B4E9'>louder</span> 
  <span style = 'color: #029E73'>higher tempo</span> 
  <span style = 'color: #E69F00'>more energetic</span>,\n
  but NOT
  <span style = 'color: #FF2500'>**more popular**</span>")
```

# Attempt 1:

```{r}

ggplot(fig2_data, aes(x=decades, y=property, fill=mean)) + geom_tile() +
  scale_fill_viridis_c(option = "A") + 
  labs(y="Property", x="Decade") + 
  theme(legend.position = "top")

```

This is not bad, but also not great. I think the suggestion that I received the most is to rank the property based on the color, so I did that. However, I still feel because I calculated the mean of song's property, though I normalized scores, the range of each property is still very different, which makes the change within some properties impossible to see.

# Attempt 2:

```{r}
ggplot(fig2_data2, aes(x=year, y=value)) + 
  geom_line(aes(color=property)) +
  gghighlight(use_direct_label = FALSE)+
  facet_wrap(~property)+
  theme(legend.position = "none")
```

I realized there is another dataset that is already ranked by year, so I thought I would try to clean that dataset and see if that works better. I also take a look at a lot more properties just to see what properties changes more or are more interesting. I end up thinking I should try to take in the characteristics that are interesting and tell a story. E.g. I found 3 increasing (energy, loudness, tempo) properties as well as 3 decreasing (acousticness, instrumentalness, speechiness) pnes. Popularity and duration are something that people are generally interested in as well. 

